\newpage
\section*{ACKNOWLEDGMENT}
\addcontentsline{toc}{chapter}{ACKNOWLEDGMENT}

My journey into this thesis began with a question I asked myself as an ICT student who loves building products: how can computer vision solve real-world problems that matter? I was not content with toy datasets or classroom exercises—I wanted to work on something where the technology tangibly helps people.

This aspiration led me to join the MycoAI lab under the guidance of \textbf{Dr. Oanh Nguyen} from Hanoi University of Science and Technology (HUST), a researcher deeply dedicated to computer vision. Through Dr. Oanh, I was introduced to \textbf{Professor Duong Vu}, a data scientist based in the Netherlands who has spent years specializing in fungal classification and founded the MycoAI laboratory to bridge AI and mycology.

Rather than handing me a well-defined problem, they gave me something more valuable: access to a real fungal image dataset and a single open-ended question—\textit{``What can you do to help a mycologist?''}

That question became my compass. I spent the early months not building models, but exploring the dataset, reading mycology literature, and speaking with domain experts to understand their workflow. Mycologists spend hours manually comparing colony photographs against reference books. A misidentification can delay research, misdirect treatment, or waste valuable lab resources. The bottleneck was not the microscope—it was the lookup.

I would like to express my sincere gratitude to \textbf{Dr. Oanh Nguyen} for her unwavering support throughout this project—for the guidance, the ideas, and the deep knowledge of computer vision techniques she shared with me at every stage. Her mentorship shaped not just the outcome of this thesis, but how I approach a problem as an engineer.

I am equally grateful to \textbf{Professor Duong Vu}, whose clear articulation of the project's requirements grounded this work in real mycologist needs. His domain expertise kept the system honest: every feature and every experiment was measured against the question of whether it actually helps the person in the lab.

\vspace{1cm}
\begin{flushright}
\textit{Hanoi, June 2025}
\end{flushright}
