\newpage
\section*{ABSTRACT}
\addcontentsline{toc}{chapter}{ABSTRACT}

This thesis presents \textbf{MycoAI Retrieval}, an image-based fungal species identification system that combines a retrieval-based classification pipeline with a production-ready web application and an Agentic Engineering development methodology.

The retrieval pipeline operates on 435 Petri dish images across 8 \textit{Penicillium} species and 31 strains. Colony images are segmented using K-Means clustering with a novel Local K=2 Shrink method to handle agar flare, then embedded with interchangeable deep and hand-crafted feature extractors. Retrieval reaches up to 81.0\% accuracy in the broader media-strategy sweep, while the canonical weighted/E1/$K=7$ comparison peaks at 71.4\% for ResNet50, 71.4\% for MobileNetV2 fine-tuned, and 54.8\% for EfficientNetB1. Five-fold cross-validation confirms 82.9\%\ mean accuracy (E2, freq\_strength, $K=3$, EfficientNetB1 finetuned). Vector search over Qdrant converts segment-level similarity into species-level predictions, with no retraining required when new species are added.

The web application wraps this pipeline into a governed, multi-role product (User / Data Owner) with 47 REST API endpoints, 14 database tables, and a dual-database architecture (PostgreSQL for governance, Qdrant for search). The system supports the complete lifecycle: upload $\to$ segment $\to$ retrieve $\to$ provide feedback $\to$ review $\to$ index $\to$ maintain model.

The development process itself was accelerated by \textbf{Agentic Engineering}: a five-agent orchestration layer (Autolab, Researcher, Planner, Worker, Reporter) that autonomously proposes hypotheses, runs experiments in isolated git worktrees, and surfaces the best-performing configurations---reducing the manual experiment cycle to an automated loop.
