\section{Conclusion}\label{chapter-5-conclusion}

\subsection{Summary}\label{summary}

This thesis presented \textbf{MycoAI Retrieval}, a complete system for fungal species identification. By combining a KNN-based retrieval model with a robust web application and an innovative Agentic Engineering workflow, the project demonstrated that AI can not only be the product but also the primary driver of the development process.

\subsection{Future Work}\label{future-work}

Future enhancements will focus on:

\begin{itemize}
\item
  \textbf{Interactive segmentation with automatic model finetuning.} The current segmentation pipeline uses classical computer vision techniques (K-Means and contour detection). When a user manually corrects a segmentation result---adjusting colony boundaries, splitting merged colonies, or excluding debris---the system should automatically finetune a YOLO-based detection model on these corrections. Over time, the segmentation model becomes personalized to the lab's imaging setup, reducing manual intervention. This closed feedback loop transforms occasional user corrections into persistent model improvements without requiring the user to understand machine learning.
\item
  \textbf{Autonomous training workflow.} Currently, training or finetuning the feature extractor requires the user to manually download the dataset, set up a training environment, and run training scripts. A future version should provide an in-application training interface: the user selects reference strains, chooses a backbone architecture, and triggers training with a single click. Training progress, validation metrics, and model checkpoints would be tracked within the web application, and the best model would be automatically promoted to the active feature extractor. This reduces the barrier to model improvement to zero---a mycologist with no ML background can update the system when new reference data becomes available.
\item
  \textbf{Batch processing and plate-level analysis.} A single laboratory session may produce dozens of Petri dishes across multiple species and media. The system should support batch upload with automatic plate assignment: the user photographs an entire incubation batch, and the system segments all plates, extracts features, and produces a structured report mapping each dish to its predicted species with confidence scores. This turns the tool from a one-at-a-time lookup into a laboratory workflow accelerator.
\item
  \textbf{Growth-medium-aware ensemble weighting.} Different growth media emphasize different morphological traits---CYA highlights colony texture while YES emphasizes pigmentation. A future iteration could learn per-medium confidence weights, dynamically adjusting which feature extractors and reference neighbors are emphasized based on the query image's growth medium. This leverages the dataset's balanced multi-media design to improve accuracy without requiring additional data.
\item
  \textbf{Integration with taxonomic databases.} Linking the system to public taxonomic resources (e.g., MycoBank, Index Fungorum) would allow automatic retrieval of species descriptions, known metabolite profiles, and taxonomic histories alongside visual predictions. This contextual information helps the mycologist interpret retrieval results and catches cases where the visual prediction conflicts with known species characteristics.
\end{itemize}
